---
title: "MTP - HR signatures interrogation"
output: html_document
date: "2025-01-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(GSVA)
library(tidyverse)
library(patchwork)
library(ggpubr)
library(presto)
```

## Importing our objects

```{r}
# seurat
luo_seu <- readRDS("~/GSE222556_processed.rds")
lau_seu <- readRDS("~/GSE266577_processed.rds")
# objects
luo <- readRDS("GSE222556_objects.rds")
lau <- readRDS("GSE266577_objects.rds")
#signatures
publ_sigs <- readRDS("publ_sigs.rds")
common_sig <- readRDS("common_sig.rds")
# str(luo)
# str(lau)
```

## Interrogating HR signatures from Luo et al., 2024

### On the same dataset

The code is explained for this example and applies to the following chunks.

```{r, fig.width=6, fig.height=6}
# Define the count matrix to interrogate
collapse_samp <- luo$tables$counts_samp
head(collapse_samp)[1:5,1:5]

# Create gsva object with the selected signatures
# Select HRD or HRP signatures from the list
collapsed_gsvapar<-gsvaParam(collapse_samp, luo$signatures$HRD)

# Apply gsva()
collapsed_gsva<-gsva(collapsed_gsvapar)
head(collapsed_gsva)[1:5,1:5]

# Long format to the matrix
collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)
head(collapsed_gsva)

# Add molecular status from the corresponding seurat
collapsed_gsva<-merge(
  collapsed_gsva,
  luo_seu@meta.data %>% dplyr::select(Sample, mol) %>% 
    unique)
head(collapsed_gsva)

# Plot
collapsed_gsva %>% 
  ggplot(aes(mol, score))+
  geom_violin(trim = FALSE)+
  geom_jitter(width = 0.1)+
  facet_wrap(.~sig, scales="free", ncol=2)+
  stat_compare_means(label.x.npc = .3, size=3, color="blue")+
  labs(x=NULL, y="GSVA score")+
  theme_bw()
```

### On Launonen et al., 2024

```{r, fig.width=6, fig.height=6}
collapse_samp <- lau$tables$counts_samp
collapsed_gsvapar<-gsvaParam(collapse_samp, luo$signatures$HRD)
collapsed_gsva<-gsva(collapsed_gsvapar)
collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)
colnames(lau_seu@meta.data)[4] <- "Sample"
collapsed_gsva<-merge(
  collapsed_gsva,
  lau_seu@meta.data %>% dplyr::select(Sample, mol) %>% 
    unique)
collapsed_gsva %>% 
  ggplot(aes(mol, score))+
  geom_violin(trim = FALSE)+
  geom_jitter(width = 0.1)+
  facet_wrap(.~sig, scales="free", ncol=2)+
  stat_compare_means(label.x.npc = .3, size=3, color="blue")+
  labs(x=NULL, y="GSVA score")+
  theme_bw()
```

## Interrogating HR signatures from Launonen et al., 2024

### On the same dataset

```{r, fig.width=6, fig.height=6}
collapse_samp <- lau$tables$counts_samp
collapsed_gsvapar<-gsvaParam(collapse_samp, lau$signatures$HRD)
collapsed_gsva<-gsva(collapsed_gsvapar)
collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)
collapsed_gsva<-merge(
  collapsed_gsva,
  lau_seu@meta.data %>% dplyr::select(Sample, mol) %>% 
    unique)
collapsed_gsva %>% 
  ggplot(aes(mol, score))+
  geom_violin(trim = FALSE)+
  geom_jitter(width = 0.1)+
  facet_wrap(.~sig, scales="free", ncol=2)+
  stat_compare_means(label.x.npc = .3, size=3, color="blue")+
  labs(x=NULL, y="GSVA score")+
  theme_bw()
```

### On Luo et al., 2024

```{r, fig.width=6, fig.height=6}
collapse_samp <- luo$tables$counts_samp
collapsed_gsvapar<-gsvaParam(collapse_samp, lau$signatures$HRD)
collapsed_gsva<-gsva(collapsed_gsvapar)
collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)
collapsed_gsva<-merge(
  collapsed_gsva,
  luo_seu@meta.data %>% dplyr::select(Sample, mol) %>% 
    unique)
collapsed_gsva %>% 
  ggplot(aes(mol, score))+
  geom_violin(trim = FALSE)+
  geom_jitter(width = 0.1)+
  facet_wrap(.~sig, scales="free", ncol=2)+
  stat_compare_means(label.x.npc = .3, size=3, color="blue")+
  labs(x=NULL, y="GSVA score")+
  theme_bw()
```

## Interrogating common signatures

We generated a list of "common" signatures combining common features between 
different signatures and datasets (see MTP_HRsig_generation)

```{r}
# Explore the signature list and the genes
common_sig
```

### On Luo et al., 2024

```{r, fig.width=6, fig.height=6}
collapse_samp <- luo$tables$counts_samp
collapsed_gsvapar<-gsvaParam(collapse_samp, common_sig)
collapsed_gsva<-gsva(collapsed_gsvapar)
collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)
collapsed_gsva<-merge(
  collapsed_gsva,
  luo_seu@meta.data %>% dplyr::select(Sample, mol) %>% unique)
collapsed_gsva %>% 
  ggplot(aes(mol, score))+
  geom_violin(trim = FALSE)+
  geom_jitter(width = 0.1)+
  facet_wrap(.~sig, scales="free", ncol=2)+
  stat_compare_means(label.x.npc = .3, size=3, color="blue")+
  labs(x=NULL, y="GSVA score")+
  theme_bw()
```

### On Launonen et al., 2024

```{r, fig.width=6, fig.height=6}
collapse_samp <- lau$tables$counts_samp
collapsed_gsvapar<-gsvaParam(collapse_samp, common_sig)
collapsed_gsva<-gsva(collapsed_gsvapar)
collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)
collapsed_gsva<-merge(
  collapsed_gsva,
  lau_seu@meta.data %>% dplyr::select(Sample, mol) %>% unique)
collapsed_gsva %>% 
  ggplot(aes(mol, score))+
  geom_violin(trim = FALSE)+
  geom_jitter(width = 0.1)+
  facet_wrap(.~sig, scales="free", ncol=2)+
  stat_compare_means(label.x.npc = .3, size=3, color="blue")+
  labs(x=NULL, y="GSVA score")+
  theme_bw()
```

## Selecting signatures for HR status prediction

Similar analyses will be performed using the count matrix from Vázquez et al.,
2022 as a validation dataset.

### Dataset preparation

Upload a downsized version of the dataset for signature interrogation

```{r}
# Object
vaz_seu <- readRDS("~/Shared_Juan/vazquez_2022_ds.rds")
# head(vaz_seu@meta.data)
# colnames( vaz_seu@meta.data)

colnames(vaz_seu@meta.data)[c(4)] <- "Sample"

# Create pseudobulk matrix
vaz_coll.samp<-presto::collapse_counts(counts_mat = vaz_seu@assays$RNA@counts, 
                                        meta_data = vaz_seu@meta.data, 
                                        varnames = c("Sample"), 
                                        how="sum")

# Add column names according to the varnames
colnames(vaz_coll.samp$counts_mat)<-vaz_coll.samp$meta_data$Sample

# Format long
vaz_coll.samp$counts_long<-vaz_coll.samp$counts_mat %>%
  as.data.frame() %>%
  rownames_to_column("Gene") %>%
  gather(Sample, score, -Gene)

# Add HR status from seurat metadata
vaz_coll.samp$counts_long<-merge(vaz_coll.samp$counts_long,
                                 vaz_seu@meta.data %>% dplyr::select(Sample, mol) %>% unique)

# Filter out "Undetermined" samples
vaz_coll.samp$counts_long<-dplyr::filter(vaz_coll.samp$counts_long, mol %in% c("HRD","HRP"))

# Define matrix
vaz_counts<-vaz_coll.samp$counts_long %>% 
  dplyr::select(Sample, Gene, score) %>% 
  spread(Sample, score) %>%
  column_to_rownames("Gene") %>% 
  as.matrix

# Define colData
vaz_colData<-vaz_seu@meta.data %>% 
  dplyr::select(Sample, mol) %>% 
  unique()

```

### Signature interrogation

```{r}
collapse_samp <- vaz_counts

# Need to modify sig names in order to interrogate all of them at once
# names(lau_obj$signatures$HRD)<-paste0("lau_HRD_",names(lau_obj$signatures$HRD))
# names(luo_obj$signatures$HRD)<-paste0("luo_HRD_",names(luo_obj$signatures$HRD))
# names(lau_obj$signatures$HRP)<-paste0("lau_HRP_",names(lau_obj$signatures$HRP))
# names(luo_obj$signatures$HRP)<-paste0("luo_HRP_",names(luo_obj$signatures$HRP))

collapsed_gsvapar<-gsvaParam(collapse_samp, c(luo_obj$signatures$HRP,luo_obj$signatures$HRD,
                                              lau_obj$signatures$HRP,lau_obj$signatures$HRD,
                                              common_sig, publ_sigs))

# collapsed_gsvapar<-gsvaParam(collapse_samp, c(common_sig, publ_sigs))

collapsed_gsva<-gsva(collapsed_gsvapar)

collapsed_gsva<-collapsed_gsva %>%
  as.data.frame() %>%
  rownames_to_column("sig") %>%
  gather(Sample, score, -sig)

collapsed_gsva<-merge(
  collapsed_gsva,
  vaz_seu@meta.data %>% dplyr::select(Sample, mol) %>% unique)

# saveRDS(collapsed_gsva, "collapsed_gsv.rds")

# Select signatures with significant p value for prediction
collapsed_gsva %>% 
  group_by(sig) %>% 
  rstatix::wilcox_test(score~mol) %>% 
  rstatix::adjust_pvalue() %>% 
  dplyr::filter(p < 0.05)

# Create a new objects with predictor signatures
sig_pred <- collapsed_gsva %>% 
  group_by(sig) %>% 
  rstatix::wilcox_test(score~mol) %>% 
  rstatix::adjust_pvalue() %>% 
  dplyr::filter(p < 0.05) %>% 
  pull(sig)

# collapsed_gsva %>%
#   dplyr::filter(sig %in% sig_pred) %>% 
#   ggplot(aes(mol, score))+
#   geom_violin(trim = FALSE)+
#   geom_jitter(width = 0.1)+
#   facet_wrap(.~sig, scales="free", ncol=3)+
#   stat_compare_means(label.x.npc = .3, size=3, color="blue")+
#   labs(x=NULL, y="GSVA score")+
#   theme_bw()

# saveRDS(sig_pred, "sig_pred.rds")
# saveRDS(sig_pred, "sig_pred2.rds")

# summary(glm(mol~Launonen_HRP+Lee+Luo_HRP+Pan+Tak_Down+Tak_Up, 
#             data=collapsed_gsva %>% spread(sig, score) %>% mutate(mol=case_when(mol == "HRD"~1,T~0))))


```