---
title: "Deciphering the immune dynamics according to homologous recombination status
in advanced ovarian cancer patients treated with neoadjuvant chemotherapy"
subtitle: "Master's thesis program" 
author: "Juan A. Marín Jiménez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R update and Seurat package installation

I followed installation instructions for Seurat_v5 at https://satijalab.org/seurat/articles/install_v5.html but needed to reinstall a 
lot of dependencies before

```{r}
if(0){
# R update needed
# Save packages
tmp <- installed.packages()
installedpkgs <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
save(installedpkgs, file="installed_old.rda")
# Install R here
# Recover packages
load("installed_old.rda")
tmp <- installed.packages()
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
install.packages(missing)
update.packages()

# Seurat installation
install.packages('Seurat')

# Other recommended packages
setRepositories(ind = 1:3, addURLs = c('https://satijalab.r-universe.dev', 'https://bnprks.r-universe.dev/'))
install.packages(c("BPCells", "presto", "glmGamPoi"))

# Install the remotes package
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
install.packages('Signac')
remotes::install_github("satijalab/seurat-data", quiet = TRUE)
remotes::install_github("satijalab/azimuth", quiet = TRUE)
remotes::install_github("satijalab/seurat-wrappers", quiet = TRUE)
}
```

# Needed libraries

```{r}
if(1){
library(Seurat)
library(tidyverse)
library(ggplot2)
library(patchwork)
}
```

# GSE dataset download

Due to lack of memory, I will manually download all the files

```{r}
if(0){
# Install GEOquery
if (!requireNamespace("GEOquery", quietly = TRUE)) {
    install.packages("GEOquery")
}
library(GEOquery)
# Dataset: Zhang_2022
# gse_id <- "GSE165897"
# gse_data <- getGEO(gse_id, GSEMatrix = TRUE)
# gse_info <- getGEOSuppFiles(gse_id)
}
```

# Seurat object creation

```{r}
if(0){
# metadata
cell_info <- data.table::fread("GSE165897/GSE165897_cellInfo_HGSOC.tsv.gz") %>% 
  column_to_rownames(var = 'cell')

# exploring cell metadata and annotation categories
cbind(table(cell_info$cell_subtype))
table(cell_info$cell_subtype, cell_info$cell_type)

# count matrix
umi_counts <- data.table::fread("GSE165897/GSE165897_UMIcounts_HGSOC.tsv.gz") %>% 
  column_to_rownames(var = 'V1')
# umi_counts[1:3,1:3]
# umi_counts_HOLD <- umi_counts

# Create my Seurat object
# seu_zhang_22 <- CreateSeuratObject(counts=umi_counts, meta.data=cell_info)
# mem.maxVSize()
# Imposible to create seurat object due to vector memory limit of 16Gb

# I split the matrix in 5000 colums fragments and put them in a list
seu_zhang_22 <- list()
for (i in 1:11){
  a1 <- i*5000-4999
  a2 <- i*5000
  if(i==11){
    a2 <- ncol(umi_counts)
  }
  umi_frag <- umi_counts[,a1:a2]
  cell_frag<- cell_info[colnames(umi_frag),]
  print(all(rownames(cell_frag)==colnames(umi_frag)))
  tmp <- CreateSeuratObject(counts=umi_frag, meta.data=cell_frag)
  seu_zhang_22[[i]] <- tmp
  tmp <- NULL
}
seu_zhang_22

# Merging all the elements from seu_zhang_22
seurat_zhang_22 <- seu_zhang_22[[1]]
for (i in 2:length(seu_zhang_22)){
  seurat_zhang_22 <- merge(seurat_zhang_22, seu_zhang_22[[i]])
}
seurat_zhang_22
Layers(seurat_zhang_22)

# Need to fuse all the generated layers using JoinLayers
seurat_zhang_22[["RNA"]] <- JoinLayers(seurat_zhang_22[["RNA"]])
seurat_zhang_22
Layers(seurat_zhang_22)
}
```

# Seurat object analysis

The objective is to develop an script to get a QC report of every Seurat object
created for the project

```{r}
seurat_zhang_22[["RNA"]]$counts[1:5,1:5]
```

